Tracing the Hercules stream around the Galaxy

### 2010-05-21 tex/hercules.tex
groups-simulations: abstract and intro tweaks

\title{Tracing the Hercules stream around the Galaxy}
\author{Jo~Bovy}%\altaffilmark{1,2}}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\email{jb2777@nyu.edu}
%\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
%  New York University, 4 Washington Place, New York, NY 10003, USA}
%\altaffiltext{2}{Correspondence should be addressed to jo.bovy@nyu.edu~.}

\begin{abstract}
It has been proposed that the Hercules stream, a group of co-moving
stars in the Solar neighborhood offset from the bulk of the velocity
distribution, is the result of resonant interactions between stars in
the outer disk and the Galactic bar. Only seen so far in the immediate
Solar neighborhood, I predict the distribution of stellar velocities
and the run of the Hercules stream as a function of location in the
Galactic disk in a simple model for the Galaxy and the bar that
produces the observed Hercules stream. I further show that the
Hercules feature is strong enough that it can be unambiguously
detected in the distribution of line-of-sight velocities in selected
directions. I quantify the most promising lines-of-sight for detection
in radial velocities; These are at $l \gtrsim 250^{\circ}$. The
predictions presented here are only weakly affected by distance
uncertainties, assumptions about the distribution function of the
stellar disk, and the details of the Galactic potential including the
effect of spiral structure; \Gaia\ and future spectroscopic surveys of
the Galactic disk will be able to robustly test the origin of the
Hercules stream and constrain the properties of the Galactic bar.
\end{abstract}

@@

The velocity distribution of nearby stars contains various large
overdensities of co-moving stars that cannot be explained as the
vestiges of the clumpy nature of star formation
\citep[\eg,][]{Bovy10a}. In particular, the Hercules moving group, a
group of stars significantly offset from the bulk of the observed
velocity distribution \citep[\eg,][see
  \figurename~\ref{fig:obs}]{Dehnen98b,Bovy10a}, displays a wide range
of ages and metallicities
\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
that has it originating through resonant interactions of stars in the
outer disk with the bar in the central region of the Galaxy has been
put forward \citep{dehnen00a,fux01a}, but any confirmation of this
picture beyond the locallly observed stellar kinematics remains
lacking. An investigation of the Hercules members' color--magnitude
diagram found no evidence for any significant metallicity anomaly with
respect to other local stars \citep{Bovy10a}, even though such a
difference should be expected as the Hercules stars originate from a
few kiloparsec toward the Galactic center where the average
metallicity is higher than in the Solar neighborhood. The absence of
the predicted metallicity anomaly could potentially be explained
through, \eg, radial mixing \citep{sellwood02a}, but it is clear that
chemodynamical modeling of the Galactic disk is currently not up to
the task of separating out these effects.

>>>>

\title{Tracing the Hercules stream around the Galaxy}
\author{Jo~Bovy}%\altaffilmark{1,2}}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\email{jb2777@nyu.edu}

\begin{abstract}
It has been proposed that the Hercules stream, a group of co-moving
stars in the Solar neighborhood offset from the bulk of the velocity
distribution, is the result of resonant interactions between stars in
the outer disk and the Galactic bar. So far it has only been seen in
the immediate Solar neighborhood. I predict the distribution of
stellar velocities and the run of the Hercules stream as a function of
location in the Galactic disk in a simple model for the Galaxy and the
bar that produces the observed Hercules stream. I further show that
the Hercules feature is strong enough to be unambiguously detected in
the distribution of line-of-sight velocities in selected directions. I
quantify the most promising lines-of-sight for detection in radial
velocities; these directions are all at $l \gtrsim 250^{\circ}$. The
predictions presented here are only weakly affected by distance
uncertainties, assumptions about the distribution function of the
stellar disk, and the details of the Galactic potential including the
effect of spiral structure. \Gaia\ and future spectroscopic surveys of
the Galactic disk will be able to robustly test the origin of the
Hercules stream and constrain the properties of the Galactic bar.
\end{abstract}

@@

The velocity distribution of nearby stars contains various large
overdensities of co-moving stars that cannot be explained as the
vestiges of the clumpy nature of star formation
\citep[\eg,][]{Bovy10a}. In particular, the Hercules moving group, a
group of stars significantly offset from the bulk of the observed
velocity distribution \citep[\eg,][see
  \figurename~\ref{fig:obs}]{Dehnen98b,Bovy10a}, displays a wide range
of ages and metallicities
\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
that has the Hercules stream originating through resonant interactions
of stars in the outer disk with the bar in the central region of the
Galaxy has been put forward \citep{dehnen00a,fux01a}, but any
confirmation of this picture beyond the locally observed stellar
kinematics remains lacking. An investigation of the Hercules members'
color--magnitude diagram found no evidence for any significant
metallicity anomaly with respect to other local stars \citep{Bovy10a},
even though such a difference should be expected as the Hercules stars
originate from a few kiloparsec toward the Galactic center where the
average metallicity is higher than in the Solar neighborhood. The
absence of the predicted metallicity anomaly could potentially be
explained, \eg, through radial mixing \citep{sellwood02a}, but it is
clear that chemodynamical modeling of the Galactic disk is currently
not up to the task of separating out these effects.
